---
title: "ECON 187 Project 2"
author: "Arif Abd Aziz, Jon Girolamo, Shanie Talor"
date: "2023-05-18"
output: pdf_document
---

#Notes
#pull 
#add/save
#commit 
#pull(to get someone elses)/push (to give yours to everyone else) 

# Brief Introduction to Part I:

The data we found explores the relationship between the fat content of multiple different food items against the incidence and prevalence of COVID 19. The data is aggregated from 170 countries (170 observations), and we plan to use a variety of non-linear techniques to create a model that predicts COVID contraction rates from aggregated nutrition variables (pertaining to food fat content). 

# Read & Clean Covid Data:
```{r}
# Clear Environment:
rm(list = ls())

# Read Data into Environment
nutrition <- read.csv("Fat_Supply_Quantity_Data.csv", sep = ",",
                      header = TRUE)
head(nutrition)

# See Variable Names
names(nutrition)

# Distribution/Histogram of Quantitative Variables: Broken into 3 Parts
# First nine variables
par(mfrow = c(3,3)) # Break this up because need plots to look clean
for (variable in names(nutrition)[1:9]) {
  if (is.numeric(nutrition[[variable]])) {
    hist(nutrition[[variable]], main = variable, xlab = paste0(variable," values"))
  }
}

# Second group of nine variables
par(mfrow = c(3,3))
for (variable in names(nutrition)[10:18]) {
  if (is.numeric(nutrition[[variable]])) {
    hist(nutrition[[variable]], main = variable, xlab = paste0(variable," values"))
  }
}

# Last Group of Nine Variables
par(mfrow = c(3,3))
for (variable in names(nutrition)[19:27]) {
  if (is.numeric(nutrition[[variable]])) {
    hist(nutrition[[variable]], main = variable, xlab = paste0(variable," values"))
  }
}

# Last Group of 5 Variables
par(mfrow = c(3,2))
for (variable in names(nutrition)[28:32]) {
  if (is.numeric(nutrition[[variable]])) {
    hist(nutrition[[variable]], main = variable, xlab = paste0(variable," values"))
  }
}
```

There are a lot of predictors in this data set, representing the fat contents of various foods. We want to model for variables that show a high degree of variation. We can pick out these predictors from their histograms. Those that are concentrated at certain values may not give us the insights into predicting the incidence of COVID-19 in different countries. For example, aquatic products and their fat contents don't exhibit strong variation. This makes sense since some countries may not have aquatic products. Though this variable may help to predict COVID 19 incidence in countries that do consume aquatic products, it may pick up dynamics in non-aquatic consuming countries that do not exist in the data. 

We also want to remove variables that may contain similar information to others, such as "Sugar Crops" and "Sugar Sweetners". Our main response variables are COVID-19 deaths and confirmed cases. These represent proportions of a country's total population.  

```{r}
# Keep Variables That We'd Like to Study 
library(dplyr)
nutrition <- nutrition %>% select(-c("Country", "Alcoholic.Beverages",
                                     "Aquatic.Products..Other", "Oilcrops",
                                     "Sugar.Crops", "Population",
                                     "Unit..all.except.Population.", "Recovered",
                                     "Active"))

# Remove Empty NA Values, Assuming Death Rate is Our Main Response Variable
removena <- which(is.na(nutrition$Deaths))
nutrition <- nutrition[-removena,]
nrow(nutrition)

# Check to See if Any Other NA Values in Main Variables:
for (variable in names(nutrition)){
  print(paste0(variable," has NA values: ",any(is.na(nutrition[[variable]]))))
} # See That Obesity and Undernourished Variables Still Have NA Values

removena_obesity <- which(is.na(nutrition$Obesity))
removena_undernourished <- which(is.na(nutrition$Undernourished))

# Remove Further NA Values
nutritiona <- nutrition[-removena_obesity,]
nutrition_cln <- nutritiona[-c(5,11,59,120,124,125,136,144),]
nrow(nutrition_cln)
nrow(nutritiona)
```

Now that our data set is largely cleaned, removing for predictors we are not interested in and adjusting for empty values, we can now look into feature selection methods. The main one we will use is the Boruta algorithm which uses random forests to pick out a subset of best predictors. We will assume that Covid-19 deaths is our response variable. 

# Feature Selection
```{r, echo = FALSE}
# Boruta Algorithm
library(Boruta)
library(randomForest)
boruta_covid <- Boruta(Deaths~.-Confirmed, data = nutrition, doTrace = 2, 
                       randomForest = TRUE)
```

```{r}
# Boruta Plot
plot(boruta_covid, main = "Boruta Algorithm Feature Selection", las = 2)
```

According to our Boruta plot, ten out of the twenty three predictors we shortlisted are significant in explaining and predicting COVID-19 deaths. Moving forward, we'll be using these 10 predictors and applying specific transformations of them (splines, polynomials) to predict COVID-19 deaths. 


# Relationship Between Response Variable & Chosen Predictors
```{r}
regression <- lm(Deaths ~. -Confirmed, data = nutrition)
```

```{r}
regression1 <- lm(Deaths ~.-Confirmed, data = nutrition)
```


```{r}

<<<<<<< HEAD
regression2 <- lm(Deaths ~.-Confirmed, data = nutrition)
=======
#pls pls pls work 
>>>>>>> 1a056aceee799ca409d351376946dad655bf1a78

```

jon
